Time stretching of the GeV emission of GRBs: \\ Fermi-LAT data vs geometrical model

### 2015-12-18 APJ/gammaRaysAPJ.tex
Added reference to Goldstein 2015

Gamma-ray bursts (GRBs) are among the most energetic events in the
Universe. Therefore, the studies of the emission mechanism and
phenomenology of GRBs may provide the new knowledge in particle
physics. Moreover, the GRBs are observed from cosmological distances
and therefore bear an imprint of the late-time evolution of the
Universe. An extensive studies of these explosions led to a number of
interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
review. {\bf In particular, there are indications that the total energy emitted in gamma-rays during a
burst is similar for the different GRBs within an order
of magnitude~\citep{postnov1999grb,Bloom:2003wy}.
One should keep in mind, though, that there is no agreement on the width of the distribution, and it is in fact model dependent. More recent results, for example, suggest that energetics of different bursts might be different~\citep{racusin2009jet, chandra2012radio}. In this paper, however, we explore the possibility that most of the
bursts have similar energetics in their rest frames to simplify analysis. When more precise energy distribution becomes available, the analysis in this paper may be improved. {\bf Also, it is generally accepted that} the
high energy radiation of GRBs follow{\bf s} the predictions of
the synchrotron radiation scenario~\citep{Wang:2013ptaw}}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }. Several
observations are related to the temporal variations of spectra. This
way, the spectral lags were found between different low energy bands {\bf \citep{cheng1995time,Wu:1999ms} }and the very-high-energy radiation was discovered to
be extended relative to the x-ray emission
\citep{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
we go further and compare the temporal extension of GRB gamma-ray
radiation in the two energy bands $\unit[100]{MeV} < E <
\unit[1]{GeV}$ and $E > \unit[1]{GeV}$ (hereafter, low and high-energy
bands) using the data from the Large Area Telescope (LAT) of the Fermi
satellite~\citep{2009ApJ...697.1071A,Ackermann:2012kna}.

@@

>>>>

Gamma-ray bursts (GRBs) are among the most energetic events in the
Universe. Therefore, the studies of the emission mechanism and
phenomenology of GRBs may provide the new knowledge in particle
physics. Moreover, the GRBs are observed from cosmological distances
and therefore bear an imprint of the late-time evolution of the
Universe. An extensive studies of these explosions led to a number of
interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
review. {\bf In particular, there are indications that the total energy emitted in gamma-rays during a
burst is similar for the different GRBs within an order
of magnitude~\citep{postnov1999grb,Bloom:2003wy,Goldstein:2015fib}.
One should keep in mind, though, that there is no agreement on the width of the distribution, and it is in fact model dependent. More recent results, for example, suggest that energetics of different bursts might be different~\citep{racusin2009jet, chandra2012radio}. In this paper, however, we explore the possibility that most of the
bursts have similar energetics in their rest frames to simplify analysis. When more precise energy distribution becomes available, the analysis in this paper may be improved. {\bf Also, it is generally accepted that} the
high energy radiation of GRBs follow{\bf s} the predictions of
the synchrotron radiation scenario~\citep{Wang:2013ptaw}}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }. Several
observations are related to the temporal variations of spectra. This
way, the spectral lags were found between different low energy bands {\bf \citep{cheng1995time,Wu:1999ms} }and the very-high-energy radiation was discovered to
be extended relative to the x-ray emission
\citep{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
we go further and compare the temporal extension of GRB gamma-ray
radiation in the two energy bands $\unit[100]{MeV} < E <
\unit[1]{GeV}$ and $E > \unit[1]{GeV}$ (hereafter, low and high-energy
bands) using the data from the Large Area Telescope (LAT) of the Fermi
satellite~\citep{2009ApJ...697.1071A,Ackermann:2012kna}.

@@

\bibitem[Goldstein et al.(2015)]{Goldstein:2015fib}
Goldstein, A., Connaughton, V., Briggs, M.-S., Burns, E. 2015, arXiv:1512.04464
